Report

### 2023-03-20 manuscript.Rmd
update 2023-03-20 14:39:27

---
title             : "Tutorial: Aggregate evidence from heterogeneous replication studies using the product Bayes factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "1"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Utrecht University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"

@@

abstract: |
  TODO

@@

bibliography      : ["product_bayes.bib"]

@@

```{r setup, include = FALSE}
run_everything <- FALSE
library("papaja")
library(tidySEM)
library(kableExtra)
library(pwr)
r_refs("r-references.bib")
out <- readRDS("out.RData")
```

@@

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has come into focus as a solution to this crisis and as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
<!--For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].-->
<!--A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations]. -->
<!--Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].-->
In step with this interest in replication research,
research synthesis methods have blossomed [REF].
These methods aggregate the findings of multiple studies,
and thus enable drawing overarching conclusions across multiple studies.
The present paper focuses on quantitative methods for research synthesis [e.g., @vanlissaSelectRelevantModerators2021], although qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods also exist.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators. 
Each of these approaches require making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

An alternative approach for evidence synthesis that does not impose assumptions about heterogeneity is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
Instead of aggregating effect sizes, BES aggregates the evidence in favor of an informative hypothesis $H_i$ across studies.
The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
A Bayes factor can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
For the purpose of this paper, we will assume that all Bayes factors are the Bayes factors of an informative hypothesis $H_i$ against its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
In this context, the subscript $_!$ represents the negation operator; in other words, $H_{!i}$ means "not $H_i$".
Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
these Bayes factors can be synthesized across studies by taking their product.
The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
The only assumption of BES is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the value or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers the question: Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the informative hypothesis?
Both methods thus provide complementary information.

This tutorial paper introduces the first implementation of BES in useruse-friendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

<!-- Rebecca: -->
<!-- ms meer over het model vertellen. -->
<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
<!-- Dus bijv: -->
<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho. We hebben k studies die elk een sample van n hebben getrokken/verzameld. Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen). Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is. En dan iets als: Dit doen we 1000 keer per set van condities.  -->
<!-- En dan nog iets over reliability ook. -->
The present simulation study set out to validate the PBF algorithm.
For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
<!-- CJ: Hier staat multiple samples; elders multiple groups / studies / etc. Kan je die terminologie gelijktrekken? -->
The informative hypothesis was kept constant at $H_i: \rho > .1$.
All simulation conditions were evaluated once in the presence of a true population effect,
defined as $\rho = .2$,
and once in the presence of a null effect, defined as $\rho = .1$.
A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported over its complement.
This threshold is subjective but rooted in conventional standards for interpreting Bayes factors [@jeffreys1998theory].
<!-- Rebecca: -->
<!-- waar is > 3 op gebaseerd? -->
<!-- Kass&Raftery? Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten). Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.  -->
<!-- Of vertel over betting odds... -->
<!-- Maar het blijft een dichitome beslissing zo.... -->
<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
<!-- Rebecca: -->
<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
<!-- In practice it may be true half of the time, because of sampling, but still. -->
<!-- This should be reflected by BF is approx 1. -->
<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The number of observations per sample $n \in (50, 200, 500, 800)$,
which (assuming a known effect of $\rho = .1$) provides statistical power to reject a false null hypothesis at $\alpha = .05$ of
$(`r paste0(formatC(pwr.r.test(c(50, 200, 500, 800), .1)$power, digits = 2, format=  "f"), collapse = ", ")`)$, respectively.
the number of samples $k \in (2, 3, 10)$, where 2 is the minimum number of samples possible and 10 is a relatively large number, considering that replications are not very common.
The reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
assumed when analyzing correlations between observed items or scale scores.
For all unique combinations of the design factors, 1000 data sets were generated.

@@

The main algorithm of interest was the PBF.
As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was vote counting: counting the number of significant effects.
Although this method is in use for aggregating conceptual replications,
it should not be considered good practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [(Scheerens, 2015):].
Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
The third algorithm was individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.
<!-- Rebecca: -->
<!-- zou je dit vooraf moeten zeggen? -->
<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->

@@

For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
This results in a so-called "confusion matrix",
which gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
These quantities were summarized as sensitivity, ${TP}{TP + FN}$, the probability of making the inferential decision that the informative hypothesis is supported, given that it was indeed true in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $\frac{TN}{TN+FP}$, the probability of concluding the informative hypothesis is not supported, given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
Recall that sensitivity is the probability of concluding that the informative hypothesis is supported, given that it was indeed true in the population,
and specificity, is the probability of concluding that it is not supported, given that it was false in the population.
The results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
In contrast, specificity was very high for all algorithms except PBF.
This suggests that the other algorithms classified most conditions as negatives (no effect found),
regardless of the existence of a population effect.
The PBF trades a loss of specificity for increased sensitivity.
At the same time, the average specificity and sensitivity were approximately equal, see Table \@ref(tab:tabconf).

@@

We examined the effect of simulation conditions on overall accuracy.
```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
<!-- Rebecca: Nu niet compleet? -->
<!-- Gek om steeds ene paragraaf van 1 zin te hebben iig. -->
<!-- Verder mist er denk ik info. Ik zou graag eerst meer uitleg zien. Hoe heb je dat examine, wat ga je laten zien (niet het resultaat an sich, snap je wat ik bedoel...). -->
<!-- Verder zou ik denk ik niet hier al weggeven wat je hierna gaat laten zien (en anders dat explicieteren dat dat is wat je doet). -->

@@

```{r fign, fig.cap="Mean performance by sample size"}
res <- readRDS("confusion_by_cond.RData")

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")
```

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of reliability on the difference between algorithms (see Table \@ref(tab:tabeffect)).

### Effect of number of groups

>>>>

---
title             : "Aggregating evidence from conceptual replication studies using the product Bayes factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "2"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Rebecca Kuiper"
    affiliation: "2"
    role:
      - Conceptualization
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Tilburg University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Utrecht University, dept. Methodology & Statistics"

@@

abstract: |
  The product Bayes factor (PBF) can synthesize evidence in favor of an informative hypothesis across heterogeneous replication studies. It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated. The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples. The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability. PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA). This was primarily due to PBF’s greater sensitivity in detecting a true effect. However, PBF had relatively lower specificity. The PBF showed increasing sensitivity and specificity with increasing sample size. With an increasing number of samples, lower sensitivity was traded for greater specificity. Although PBF's overall performance was less susceptible to reliability than the other algorithms, this masked a trade-off between reliability and specificity. PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies. Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.

@@

bibliography      : ["product_bayes.bib", "references.bib"]

@@

```{r setup, include = FALSE}
run_everything <- TRUE
library("papaja")
library(tidySEM)
library(kableExtra)
library(pwr)
library(bain)
r_refs("r-references.bib")
source("UI function/pbf.R")
out <- readRDS("out.RData")
```

@@

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
In step with this interest in replication research,
research synthesis methods have become increasingly popular.
These methods aggregate research findings,
and thus enable drawing overarching conclusions across multiple (replication) studies.
In this paper, we introduce a new research synthesis method that uses Bayesian evidence synthesis to aggregate the evidence in favor of an informative hypothesis, quantified by the Bayes factor,
across multiple studies.
This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
We describe the method in detail and demonstrate its application through a tutorial example analysis.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
When studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations, these between-study differences can introduce heterogeneity in findings.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications,
one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be random,
random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectRelevantModerators2021].
However, accounting for moderators requires a relatively high number of observations per moderator,
which may not be available.
Each of these approaches thus makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
but are otherwise very heterogeneous.
The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
<!-- even machine learning based methods cannot identify relevant moderators. -->
It is still possible, however, to quantify the support these studs provide for the underlying informative hypothesis.
To this end, Bayesian evidence synthesis (BES) aggregates the evidence in favor of an informative hypothesis $H_i$ across studies,
without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].

The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
The BF can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
Within the scope of this paper,
all Bayes factors are the ratio of evidence in favor of an informative hypothesis $H_i$ relative to its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
The subscript $_!$ represents the negation operator;
in other words, $H_{!i}$ means "not $H_i$".
Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
A value of $BF_c = 10$ means that the data provide ten times more support in favor of the hypothesis than against it.

When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
these Bayes factors can be synthesized across studies by taking their product [@kuiperCombiningStatisticalEvidence2013].
The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
The only assumption of the PBF is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
Note that other approaches to BES exist;
for instance, it is possible to use the posterior of one study as the prior for a replication study, and thus accumulate evidence across studies [see @heckReviewApplicationsBayes2022].
Such applications are out of scope of the present paper,
which addresses the PBF approach to BES.

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the point estimate or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers questions like:
Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the hypothesis of interest?
Both methods are appropriate for different research questions, and provide complementary information.

This paper introduces the first implementation of BES in user-friendly free open source software (FOSS).
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `0.2.9`.
<!-- `r as.character(packageVersion("bain"))`. -->
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

<!-- Rebecca: -->
<!-- ms meer over het model vertellen. -->
<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
<!-- Dus bijv: -->
<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho. We hebben k studies die elk een sample van n hebben getrokken/verzameld. Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen). Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is. En dan iets als: Dit doen we 1000 keer per set van condities.  -->
<!-- En dan nog iets over reliability ook. -->
The present simulation study set out to validate the PBF algorithm and benchmark it against other evidence synthesis methods.
We simulated a scenario where an informative hypothesis about a correlation between two variables was measured across several independent samples,
and the resulting evidence was synthesized across samples using multiple methods.
<!-- Rebecca: -->
<!-- waar is > 3 op gebaseerd? -->
<!-- Kass&Raftery? Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten). Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.  -->
<!-- Of vertel over betting odds... -->
<!-- Maar het blijft een dichitome beslissing zo.... -->
<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
<!-- Rebecca: -->
<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
<!-- In practice it may be true half of the time, because of sampling, but still. -->
<!-- This should be reflected by BF is approx 1. -->
<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
The informative hypothesis was kept constant at $H_i: \rho > .1$.
To examine the performance of the different evidence synthesis methods in a range of scenarios,
several design factors were manipulated.
First was the presence or absence of a true population effect.
A present effect was defined as $\rho = .2$;
a null effect was defined as $\rho = .1$.
The second design factor was the number of observations per sample $n \in (50, 200, 500, 800)$.
These values were chosen because they correspond to a statistical
<!-- pwr.r.test(c(50, 200, 500, 800), .1) -->
power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
assuming $\alpha = .05$ and a known effect size of $\rho = .1$.
Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$,
where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
assumed when analyzing correlations between observed items or scale scores.
For all unique combinations of these design factors,
the simulation was repeated 1000 times.

@@

The main algorithm of interest was the PBF.
As a decision criterion to conclude that $H_i$ was supported over its complement,
we used $PBF > 3$ - a conventional threshold for inference using Bayes factors [@jeffreys1998theory].

As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was *vote counting*: counting the number of significant effects.
Although this method is still in use for aggregating conceptual replications,
it is considered bad practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [@hedgesVotecountingMethodsResearch1980].
Our vote counting algorithm summed the number of one-sided z-tests of a null hypothesis corresponding to the informative hypothesis,
so $H_0: \rho = .1$ and $H_a: \rho > .1$, which corresponds to $H_i$.
The decision criterion was that the hypothesis was supported in the majority of samples.
Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).

The third algorithm was *random-effects meta-analysis* (RMA),
which is the current gold standard for evidence synthesis [@viechtbauerConductingMetaanalysesMetafor2010].
For this algorithm, the null-hypothesis was rejected if the lower bound of a 90% confidence interval for the overall effect size excluded $H_i$.
Note that this corresponds to a test at $\alpha = .05$,
because all effects in the simulation are directional.

The fourth algorithm was *individual participant data* (IPD) meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
Like classic meta-analysis (RMA), IPD is a multilevel model, clustered by sample.
IPD uses the raw data, which makes it possible to estimate variance at the first level.
By contrast, RMA treats the first-level variance as known.
Note that the PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
With this in mind, it is informative to benchmark it against both of these methods.
IPD was also evaluated using the lower bound of a 90% confidence interval for the overall effect size.
<!-- Rebecca: -->
<!-- zou je dit vooraf moeten zeggen? -->
<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->

@@

For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
the ability to detect an effect given that it was indeed true in the population,
<!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $\frac{TN}{TN+FP}$,
the ability to correctly conclude that the informative hypothesis is not supported,
given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
PBF had a higher overall accuracy than other algorithms
followed by IPD, then RMA and finally VC,
see Table \@ref(tab:tabconf).
This higher accuracy was primarily driven by PBF's greater sensitivity to detect a true effect compared to other algorithms.
However, PBF had a lower specificity compared to all other algorithms.
This suggests that the PBF trades a loss of specificity for increased sensitivity.

@@

```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
We used ANOVAs to examine the effect of simulation conditions on overall accuracy.
The differences between algorithms were analyzed in analyses that included two-way interactions between design factors and algorithm.
As the sample size was very large,
significance tests were uninformative.
We thus focused on interpreting the effect size of the effects of design factors.
The performance of PBF was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
The differences in the effects of sample size and number of groups were relatively small between PBF and the two best practice algorithms, RMA and IPD -
but substantial between PBF and the suboptimal VC algorithm.
The reverse pattern occurred for reliability, however:
it showed a substantial difference in effect between PBF and the two best practice algorithms only.

@@

```{r}
res <- readRDS("confusion_by_cond.RData")

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
library(ggplot2)
p <- ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  scale_linetype(guide = guide_legend(reverse = TRUE) )+
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")+
  labs(shape = "Algorithm")
ggsave("sample_size.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
```
```{r fign, fig.cap="Mean performance by sample size"}
knitr::include_graphics("sample_size.pdf")
```

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increased with sample size.
The other algorithms showed only increasing sensitivity; 
specificity was limited by a ceiling effect.
This difference explains the effect of sample size on the difference between algorithms (see Table \@ref(tab:tabeffect)).

### Effect of number of samples

```{r}
df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("k", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
p <- ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  scale_linetype(guide = guide_legend(reverse = TRUE) )+
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")+
  labs(shape = "Algorithm")
ggsave("groups.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
```
```{r figk, fig.cap="Mean performance by number of groups"}
knitr::include_graphics("groups.pdf")
```

The figure indicates that, for PBF at higher levels of $k$,
lower sensitivity was traded for greater specificity.
The other algorithms did not show this pattern, as their specificity was at a ceiling.
This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC showed decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].
